% This is the file imapptdoc.tex.  Typeset with plain TeX.
\magnification=\magstephalf
\def\beginverbatim{\par\begingroup\setupverbatim\doverbatim}
{\catcode`\|=0 \catcode`\\=12 % | is temporary escape character
  |obeylines|gdef|doverbatim^^M#1\endverbatim{#1|endgroup}}
\def\setupverbatim{\tentt \obeylines \uncatcodespecials \obeyspaces}
{\obeyspaces\global\let =\ } % let active space = control space
\def\uncatcodespecials{\def\do##1{\catcode`##1=12}\dospecials}
{\catcode`\^^M=13 \gdef\gobblecr{\ifnextchar
{\gobble}{\ignorespaces}}}
{\catcode`\ =\active\gdef\vobeyspaces{\catcode`\ \active \let \xobeysp}}
 \def\xobeysp{\leavevmode{} }
\begingroup \catcode `|=0 \catcode `[= 1
\catcode`]=2 \catcode `\{=12 \catcode `\}=12
\catcode`\\=12 |gdef|@xverbatim#1\end{verbatim}[#1|end[verbatim]]
|gdef|@sxverbatim#1\end{verbatim*}[#1|end[verbatim*]]
|endgroup
\def\makeother#1{\catcode`#112\relax}
\def\verb{\begingroup \tt \uncatcodespecials
\averb}
\def\sverb#1{\def\tempa ##1#1{##1\endgroup}\tempa}
\def\averb{\obeyspaces \frenchspacing \sverb}
\font\textfontii = cmsy10
\font\eightpt = cmr8
\def\heading#1{\medskip\noindent{\bf #1.\ }}
\def\AmSTeX{{\textfontii A}\kern-.1667em\lower.5ex\hbox
 {\textfontii M}\kern-.125em{\textfontii S}-\TeX}
\def\qed{\ifhmode\unskip\nobreak\fi\ifmmode\ifinner\else\hskip5pt\fi\fi
 \hbox{\hskip5 pt \lower 1.5 pt\hbox{\vrule width .2 pt 
 \vbox{\hrule width 4 pt height .2 pt \vskip 7.1 pt
 \hrule width 4 pt height .2 pt }\unskip\vrule width .2 pt}}}
\centerline{\bf USING  THE IMA PREPRINT STYLE FILE}
\medskip
{\eightpt\centerline{DOUGLAS N. ARNOLD and BRADLEY J. LUCIER}}
\bigskip
This document explains the use of the IMA preprint style file
to typeset preprints and reports with 
\AmSTeX. The IMA preprint style file, which is a
modification of the amsppt style of Michael Spivak, is
intended as a general purpose preprint style for mathematical papers.
It has been used by the Institute
for Mathematics and its Applications since 1987 for its preprint series.
In these instructions we assume familiarity with \AmSTeX\ and amsppt,
as documented in
{\it The Joy of \TeX\ }
by Michael Spivak.

Most of the points introduced below are illustrated in
the nonsense paper {\it A sample paper to illustrate the
IMA preprint style}.

\heading{Document style}  Include the commands \verb"\input amstex"
and \verb"\documentstyle{imappt}" near the top of your document.

\heading{Magnification} To obtain reasonable results, you {\it must} set
\verb"\magnification=\magstep 1" {\it before} the command \verb"\input
amstex".

\heading{Title} Use \verb"\title" and \verb"\endtitle".
Type the title in all caps.
Attach any footnotes (see below) that apply to the paper
as a whole to the title.

\heading{Author} Use \verb"\author" and \verb"\endauthor".
Type the author's name in all caps.  For multiple authors type the word
``and'' in lowercase.

\heading{Author's affiliation}  Give the  affiliation of each author
in a footnote (see below) attached to each author's name.  Statements
acknowledging support should be contained in the same footnote.
Do {\it not} use the \verb"\affil" or \verb"\address" macros.

\heading{Abstract} If there is an abstract, use \verb"\abstract".

\heading{Keywords} If there are keywords,
use \verb"\keywords" with commas between keywords
but without ending punctuation.

\heading{Subject classifications} If there are AMS
subject classifications, use \verb"\subjclass"
with commas between classifications
but without ending punctuation.

\heading{Footnotes} Use \verb"\footnote".
For the topmatter, use footnote symbols.
In the rest of the paper
use numbered footnotes. 
The order of footnote symbols is star (*\thinspace=\thinspace\verb"*"),
dagger (\dag\thinspace=\thinspace\verb"\dag"), double-dagger
(\ddag\thinspace=\thinspace\verb"\ddag"),
section-marker (\S\thinspace=\thinspace\verb"\S").

\heading{Headings} Use \verb"\heading" without ending punctuation.
The argument consists of the section number followed by a period
and then the section name with only the first word capitalized.
Very long papers could use \verb"\heading" and \verb"\endheading"
for major headings and \verb"\heading" for minor headings.  If
\verb"\heading" is used start heading name with a section mark
(\S\thinspace=\thinspace\verb"\S") and the section number followed by a period and the
section title with the first letter of each major word capitalized.

\heading{Equation numbers}  Use \verb"\tag".

\heading{Theorems, etc} Use \verb"\proclaim" and \verb"\endproclaim"
for theorems, lemmas, corollaries, claims, propositions, etc.  The
argument to \verb"\proclaim" takes no ending punctuation.

\heading{Definitions} Use \verb"\proclaim" and \verb"\endproclaim",
but use \verb"\rm" to set the body of the definition in roman.

\heading{Proofs, etc} Use \verb"\demo" and \verb"\enddemo"
for proofs, examples, remarks, cases, etc.  The
argument to \verb"\demo" takes no ending punctuation.  The
end-of-proof mark is an open box (\qed\ =\thinspace\verb"\qed").

\heading{Lists and sublists}  Use \verb"\roster" and \verb"\endroster".
Rosters use default labels of the form (1), (2), etc.
Subrosters are permitted.  Each subroster {\it must} be enclosed
in a separate pair of braces. Note that subrosters are {\it not}
standard \AmSTeX.
 
\heading{References} These can be done exactly as in {\it The Joy
of \TeX} using \verb"\Refs" and \verb"\ref".

\heading{Table of contents} There are no special macros provided to set
a table of contents.  If a table of contents is desired, follow these
guidelines.  Switch to eight point type (\verb"\eightpoint").
Center the word ``CONTENTS'' in caps.  Capitalize only the first word
of entries.  Runover lines align with the start of entries.
Place the section number flush left with one em to section title,
spaced dots from end of entry to page number, and two ems between the
dots and page number.  Leave 18--20 points 
of vertical space, (baseline
to baseline, before magnification) to the text above and below.

\heading{Tables} There are no special macros provided to set
tables.  Follow these guidelines.
Switch to eight point type (\verb"\eightpoint").
Center the table on the page width.  Put a minimum
of  18 points
space (baseline to baseline, before magnification) above and/or below the table.
Center the label over the table.  The label
should be of the form \verb"{\smc Table~1.}"  The title of the table should
be in slanted roman centered over the table with only the first word capitalized.
Column headings may be in set in six point type if necessary to fit properly
over the columns.  Place  standard \TeX\ rules (0.4 point thick, before
magnification)
above and below the column headings, and at the end of the table.

\heading{Figure legends} Use \verb"\topspace" or \verb"\midspace".
Set the caption as, for example,
\verb"\topspace{3 in}\caption{\eightpoint{\smc Fig\. 1.\enspace}\it Type caption here.}".

\heading{New control sequences available in the IMA preprint style}
The amount that each roster is indented is controlled by the value of
\verb"\rosterindent".  For most lists the default value need not
be changed.
To get a footer that indicates
that the paper is a draft, give the command \verb"\draft" immediately
after the \verb"\documentstyle" command.  This will also print the
date and the file name in the footer.
Should you ever want to set the footer text explicitly,
you may redefine  \verb"\footertext".

\heading{Control sequences available
in the amsppt style but not the IMA preprint style}
Do not use the commands
\verb"\affil", \verb"\endaffil", \verb"\address", \verb"\date", and
\verb"\thanks".  They have been redefined in the IMA preprint style file
to give error messages when used, because
their function has been taken over by other constructions, mainly footnotes
to the title of the paper and the authors' names.

\heading{Fonts} The font \verb"cmcsc10" is
used at 8 point (\verb"cmcsc10 at 8 pt") so there is a
caps-and-small-caps font for use in references, etc.  You must
have the appropriate \verb"pxl", \verb"pk" or \verb"gf" file for
your printer driver. (See your local \TeX\ guru.  He or she may
be able to generate the font in the right size using Metafont.) 
Alternatively, you can redefine the font \verb"\eightsmc"
to be another font which you have available.
If you have no eight point caps-and-small-caps font, then
give the command \verb"\font\eightsmc=cmr8" to substitute
eight point roman.

The file uses the newer \verb"cm" (computer modern) family of fonts
(\verb"cmr9", \verb"cmr8", \verb"cmr6", \verb"cmmi9", \verb"cmmi8",
\verb"cmmi6", \verb"cmsy9", \verb"cmsy8", \verb"cmsy6", \verb"cmbx9",
\verb"cmbx8", \verb"cmbx6", \verb"cmti9", \verb"cmti8", \verb"cmsl9",
\verb"cmsl8", \verb"cmcsc10", and \verb"cmcsc10 at 8 pt"), so you will
have to change these references to the appropriate \verb"am" (almost
modern) fonts if that is all you have available at your site.  The
style file also assumes the existence of the American Mathematical
Society fonts \verb"msxm9", \verb"msxm8", \verb"msxm6", \verb"msym9",
\verb"msym8", and \verb"msym6".  If you do not have these fonts, then
comment out the lines in imappt.sty that reference them.

\bye
